---
title: "00_Prin_Axis_"
author: "Mohammed Alrezq"
date: "2023-07-31"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction 
- This section involves working on Principal Axis Factoring (EFA) as follow:
 - EFA is run for each category of the DVs based on the conceptual framework 
 - EFA is run for all IVs in one model 

# 1. Preliminary 
## 1.1 Load libraires 
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(psych)# use perform EFA
library(purrr)
library(ggplot2)
library(gridExtra)# to arrnge plot per row and col
```
## 1.2 load data 
```{r}
model_00 <- read.csv("mean_data.csv")

# Remove Response.ID and col X from the data 
model_00 <- model_00 |> 
  select(c(-Response.ID, -X))

# Convert all integer columns to numeric
model_00 <- model_00 |> 
  mutate_if(is.integer, as.numeric)

```

# 2 Statistical descriptive
```{r}
describe(model_00)
```

# 3 Principle Axis Factoring 
## 3.2 Kaiser-Meyer-Olkin (KMO)
 
```{r}
KMO(model_00)
  #Overall MSA =  0.86 Sample adequacy. 0.60 and above is acceptable   
```

## 3.3 Bartlet’s test of sphericity
```{r}
cortest.bartlett(model_00)
#p.value = 5.660815e-161, sig. p-value and sig. of correlation between items 
```

## 3.3 Correlation 
```{r}
mode_00_correlation <- cor(model_00)
#write.csv(mode_00_correlation, "modle_00_correlation.csv")
```

# 4 SPlit data to DVs and IVs 
```{r}

# Dependent varaibes
model_00_DVs <- model_00 |> 
  select(1:13)
# Independent variables 
model_00_IVs <- model_00 |> 
  select(14:42)
```

# 5 Principal Axis Factoring for DVs 
## 5.1 determine number of factor for DVs 
```{r}
scree(model_00_DVs)
# scree recommends 1 factor 
```

## 5.2 Run Factor analysis
### Factor = 3 
```{r}
Prn_ax_DV_3 <- fa(model_00_DVs, nfactors = 3, fm = "pa", rotate = "oblimin")
print(Prn_ax_DV_3 , cut = .3, digits = 2)
```
### Factor = 2
```{r}
Prn_ax_DV_2 <- fa(model_00_DVs, nfactors = 2, fm = "pa", rotate = "oblimin")
print(Prn_ax_DV_2 , cut = .3, digits = 2)
# maximum iteration exceeded
```


## 5.2 determine number of factor for IVs 
```{r}
scree(model_00_IVs)
# scree recommends 1 factor 
```

### Factor = 3
```{r}
Prn_ax_IDV_3 <- fa(model_00_IVs, nfactors = 3, fm = "pa", rotate = "oblimin")
print(Prn_ax_IDV_3 , cut = .3, digits = 2)
```

### Factor = 4
```{r}
Prn_ax_IDV_4 <- fa(model_00_IVs, nfactors = 4, fm = "pa", rotate = "oblimin")
print(Prn_ax_IDV_4 , cut = .3, digits = 2)
```