---
title: "00_Prin_Axis_"
author: "Mohammed Alrezq"
date: "2023-07-31"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction 
- This section involves working on Principal Axis Factoring (EFA) as follow:
 - EFA is run for each category of the DVs based on the conceptual framework 
 - EFA is run for all IVs in one model 

# 1. Preliminary 
## 1.1 Load libraires 
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(psych)# use perform EFA
library(purrr)
library(ggplot2)
library(gridExtra)# to arrnge plot per row and col
```
## 1.2 load data 
```{r}
model_00 <- read.csv("mean_data.csv")

# Remove Response.ID and col X from the data 
model_00 <- model_00 |> 
  select(c(-Response.ID, -X))

# Convert all integer columns to numeric
model_00 <- model_00 |> 
  mutate_if(is.integer, as.numeric)
str(model_00)
dim(model_00)
```

# 2 Statistical descriptive
```{r}
describe(model_00)
```

# 3 Principle Axis Factoring 
## 3.1 Multivariate Normality  
```{r}
mardia(model_00)
# check what test used for normality check 
# In our case, kurtosis = -3.2 (P < 0.05). At the start and the end, the dots are away from the line. Thus, the data are not normally distributed at multivariate level. Our extraction method PAF can deal with this non-normality.
```

## 3.2 Kaiser-Meyer-Olkin (KMO)
 
```{r}
KMO(model_00)
  #Overall MSA =  0.86 Sample adequacy. 0.60 and above is acceptable   
```

## 3.3 Bartlet’s test of sphericity
```{r}
cortest.bartlett(model_00)
#p.value = 5.660815e-161, sig. p-value and sig. of correlation between items 
```

# 4 Plit data to DVs and IVs 
```{r}
names(model_00)
# Dependent variables: General success  
model_00_GEN_SUCC <- model_00 |> 
  select(1, 3, 9, 10)
# Dependent variables: LP/LSS Embeddedness   
model_00_Embd <- model_00 |> 
  select(2, 6, 7)
# Dependent variables: Improved employee results   
model_00_EMP_RES <- model_00 |> 
  select(5, 11, 12)
# Dependent variables: Improved organization results   
model_00_ORG_RES <- model_00 |> 
  select(8, 4, 13)

# Dependent varaibes
model_00_DVs <- model_00 |> 
  select(1:13)
# Independent variables 
model_00_IVs <- model_00 |> 
  select(14:42)
```

# 5 Principal Axis Factoring for DVs 
## 5.1 determine number of factor for DVs 
```{r}
scree(model_00_DVs)
# scree recommends 1 factor 
```

## 5.2 Run Factor analysis
### Factor = 4
```{r}
Prn_ax_DV_4 <- fa(model_00_DVs, nfactors = 4, fm = "pa", rotate = "oblimin")
print(Prn_ax_DV_4 , cut = .3, digits = 2)
# maximum iteration exceeded
```

### Factor = 4 and increae number of iteration = 1000
```{r}
Prn_ax_DV_4.1 <- fa(model_00_DVs, nfactors = 4, fm = "pa", rotate = "oblimin", max.iter = 1000)
print(Prn_ax_DV_4.1 , cut = .3, digits = 2)
# no improvementby maximizing iteration 
```

### Factor = 3 
```{r}
Prn_ax_DV_3 <- fa(model_00_DVs, nfactors = 3, fm = "pa", rotate = "oblimin")
print(Prn_ax_DV_3 , cut = .3, digits = 2)
```

### Factor = 3 and increae number of iteration = 1000
```{r}
Prn_ax_DV_3.1 <- fa(model_00_DVs, nfactors = 3, fm = "pa", rotate = "oblimin", max.iter = 1000)
print(Prn_ax_DV_3.1 , cut = .3, digits = 2)
# no improvementby maximizing iteration 
```

### Factor = 2 
```{r}
Prn_ax_DV_2 <- fa(model_00_DVs, nfactors = 2, fm = "pa", rotate = "oblimin")
print(Prn_ax_DV_2 , cut = .3, digits = 2)
 
```

### Factor = 5 
```{r}
Prn_ax_DV_5 <- fa(model_00_DVs, nfactors = 5, fm = "pa", rotate = "oblimin")
print(Prn_ax_DV_5 , cut = .3, digits = 2)
 

```
