---
title: "11_Prin_Axis_DVs"
author: "Mohammed Alrezq"
date: "2023-08-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Preliminary 
### 1.1 Load libraires: this need to be 
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(psych)# use perform EFA
library(purrr)
library(ggplot2)
library(gridExtra)# to arrnge plot per row and col
library(corrr)
```


```{r,}
plotTheme <- function() {
  theme(
    plot.title = element_text(size = 14, family = "sans", face = "plain", hjust = 0),
    plot.subtitle=element_text(size = 11, family = "sans", hjust = 0),
    plot.caption=element_text(size = 10, family = "sans", face = "italic", hjust = 0), 
    axis.title.x = element_text(size = 10, family = "sans", face = "plain", hjust = 1, vjust = -0.5),
    axis.title.y = element_text(size = 10, family = "sans", face = "plain", hjust = 1, vjust = 1),
    axis.text = element_text(size = 9, family = "sans", face = "plain"),
    panel.background = element_blank(),
    panel.grid.minor = element_line(colour = "gray"),
    panel.grid.major = element_line(colour = "gray"),
    axis.ticks = element_blank(),
    legend.title = element_text(size = 10, family = "sans"),
    legend.text = element_text(size = 9, family = "sans"),
    axis.line = element_blank()
  )
}
```

## 1.2 load data
```{r}
model_11 <- read.csv("mean_data.csv")

# Remove Response.ID and col X from the data 
model_11 <- model_11 |> 
  select(c(-Response.ID, -X))

# Convert all integer columns to numeric
model_11 <- model_11 |> 
  mutate_if(is.integer, as.numeric)
```

## 1.3 Split data to dependent and independet variables 
```{r}
# Dependent variables/items 
model_11_DVs <- model_11 |> 
  select(1:13)
#str(model_11_DVs)
# Independent variables/items 
model_11_IVs <- model_11 |> 
  select(14:42)
#str(model_11_IVs)
```

* Run Factor with determined factor number for *DVs *
```{r}
# with factor = 2
set.seed(1245)
model_11_DV_F2 <- psych::fa(model_11_DVs,nfactors = 2, fm = "pa", rotate = "oblimin", max.iter = 100)
psych::print.psych(model_11_DV_F2, cut = 0.3, sort=TRUE)

# Excluding CULT_INTG_5
set.seed(12333)
model_11_DV_2F_delt_CULT_INTG_5 <- psych::fa(model_11_DVs[, -which(names(model_11_DVs) == "CULT_INTG_5")],nfactors = 2, fm = "pa", rotate = "oblimin", max.iter = 100)
psych::print.psych(model_11_DV_2F_delt_CULT_INTG_5, cut = 0.3, sort=TRUE)
```


```{r}
# with factor = 3
set.seed(1234)
model_11_DV_F3 <- psych::fa(model_11_DVs,nfactors = 3, fm = "pa", rotate = "oblimin", max.iter = 100)
psych::print.psych(model_11_DV_F3, cut = 0.3, sort=TRUE)
names(model_11_DVs)

################Experimenting removing some items from DVs ###########################################
# Excluding CULT_INTG_5
set.seed(12324)
model_11_DV_CULT_INTG_5 <- psych::fa(model_11_DVs[, -which(names(model_11_DVs) == "CULT_INTG_5")],nfactors = 3, fm = "pa", rotate = "promax", max.iter = 100)
psych::print.psych(model_11_DV_CULT_INTG_5, cut = 0.3, sort=TRUE)
psych::print.psych(model_11_DV_CULT_INTG_5, sort = TRUE) # print loading w/o supressed low loadings 
# Note: this is best soltuion with 3 factros. crossloading <0.75, explained variance = 0.69, and item 10's commonality =0.38
mean(model_11_DV_CULT_INTG_5$communality)

# To download factor loading and communality for DVS
DV_loading <- model_11_DV_CULT_INTG_5$loadings
DV_communality <- model_11_DV_CULT_INTG_5$communality

DV_combined_loading_communality <- cbind(DV_loading, Communalities = DV_communality)
#write.csv(DV_combined_loading_communality, "DV_factor_loading.csv", row.names = TRUE)

# Excluding EMPL_BEL_10
set.seed(11324)
model_11_DV_EMPL_BEL_10 <- psych::fa(model_11_DVs[, -which(names(model_11_DVs) == "EMPL_BEL_10")],nfactors = 3, fm = "pa", rotate = "oblimin", max.iter = 100)
psych::print.psych(model_11_DV_EMPL_BEL_10, cut = 0.3, sort=TRUE)

# Excluding EMPL_BEL_10 and CULT_INTG_5
set.seed(11114)
model_11_DV_EMPL_BEL_10_CULT_INTG_5 <- psych::fa(model_11_DVs[, !names(model_11_DVs) %in% c("CULT_INTG_5", "EMPL_BEL_10")],nfactors = 3, fm = "pa", rotate = "oblimin", max.iter = 100)
psych::print.psych(model_11_DV_EMPL_BEL_10_CULT_INTG_5, cut = 0.3, sort=TRUE)



# Excluding EMPL_BEL_10 and CULT_INTG_5. further remove "ACT_INTG_7"
names(model_11_DVs)
set.seed(10014)
model_11_DV_delt_3items <- psych::fa(model_11_DVs[, !names(model_11_DVs) %in% c("CULT_INTG_5", "EMPL_BEL_10", "ACT_INTG_7")],
                                                 nfactors = 3, fm = "pa", rotate = "oblimin", max.iter = 100)
psych::print.psych(model_11_DV_delt_3items, cut = 0.3, sort=TRUE)

```


```{r}
# with factor = 4
set.seed(12454)
model_11_DV_F4 <- psych::fa(model_11_DVs,nfactors = 4, fm = "pa", rotate = "oblimin", max.iter = 100)
psych::print.psych(model_11_DV_F4, cut = 0.3, sort=TRUE)
```

- *Note for DVs*
 + with 5 factors, one item loaded on one factor
 + with 6 factors, Heywood case was detected
 + Best number ranged between 2 - 4
 
 
 
## Principle Component analysis 
- Determine number of factors using eigenvalues 
```{r}
# NO. of factor for DVs
model_11_DVs_PCA <- psych::principal(model_11_DVs, nfactors = length(model_11_DVs), rotate = "none")
# 2 Components > 1
```

##### PCA for DVs wiht 2 component
```{r}
# No. of factors 2 
set.seed(1114)
model_11_DVs_2PCA <- psych::principal(model_11_DVs, nfactors = 2, rotate = "oblimin")
psych::print.psych(model_11_DVs_2PCA, cut = 0.3, sort=TRUE)
```

##### PCA for DVs wiht 3 component
```{r}
# No. of factors 3
set.seed(11149)
model_11_DVs_3PCA <- psych::principal(model_11_DVs, nfactors = 3, rotate = "oblimin")
psych::print.psych(model_11_DVs_3PCA, cut = 0.3, sort=TRUE)


# Remove CULT_INTG_5
set.seed(111490)
model_11_DVs_3PCA_dlt_1item <- psych::principal(model_11_DVs[, -which(names(model_11_DVs) == "CULT_INTG_5")], nfactors = 3, rotate = "oblimin")
psych::print.psych(model_11_DVs_3PCA_dlt_1item, cut = 0.3, sort=TRUE)
```


Check Cronbach’s Alpha 
```{r}
# three latent factors for DVs based on this model:model_11_DV_CULT_INTG_5
names(model_11_DVs)
latent_factor_1 <- alpha(model_11_DVs[,c(1,4,8:13)]) # 0.94

latent_factor_2 <- alpha(model_11_DVs[,2:3]) # 0.82

latent_factor_3 <- alpha(model_11_DVs[,6:7]) # 0.82

```

