---
title: "Demographic Analysis"
author: "Mohammed Alrezq"
date: "2023-12-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(psych)
```

```{r}
All_Var_data <- read.csv("Demographic_data.csv")
All_Var_data <- All_Var_data |> 
  select(-1)

names(All_Var_data)
```

# Select only Demographic variables 
```{r}
demo_data <- All_Var_data |> 
  select(49:92)
names(demo_data)
```


```{r}
# Functional area variable 
# 1- Social services, 2- Policing force, 3- Transportation, etc
names(demo_data)
table(demo_data$FUNC_AREA)
table(demo_data$Other_Fnction)


# to comibed other with the original column Functionla area:
demo_data <- demo_data|> 
  mutate(FuncationalArea = if_else(FUNC_AREA == 11, Other_Fnction, as.character(FUNC_AREA))
)


# Suppose the first two columns are A and B, and you want X to be the third column
demo_data <- demo_data %>% select(ORG_CTRY, GOV_LVL, FUNC_AREA, Other_Fnction, FuncationalArea, everything())

# Remove extra column FUNC_AREA, and Other_Fnction
demo_data <- demo_data |> 
  select(-3:-4)

```

```{r}
# institutionalization level of LP/LSS variable 
# 1- Within a single sub-unit of the organization, 2- Within multiple sub-units of the organization, 3- Across all sub-units of the organization, 4- Other (Please specify)
names(demo_data)
table(demo_data$INST_LVL)
table(demo_data$Other_InstLvl)


# to comibed other with the original column Functionla area:
demo_data <- demo_data|> 
  mutate(InstitutionalizationLevel = if_else(INST_LVL == 4, Other_InstLvl, as.character(INST_LVL))
)


# Suppose the first two columns are A and B, and you want X to be the third column
demo_data <- demo_data %>% select(ORG_CTRY, GOV_LVL, FuncationalArea, POP_SIZE, INST_LVL, Other_InstLvl, InstitutionalizationLevel, everything())

# Remove extra column 
demo_data <- demo_data |> 
  select(-5:-6)

```

```{r}
#improvement methodology used 
#1- Lean Production, 2- Lean Six Sigma, 3- Lean Production with other general improvement methodology, 4- Other (Please specify)
names(demo_data)
table(demo_data$IMP_FOCUS)
table(demo_data$Other_ManFocus)


# to comibed other with the original column Functionla area:
demo_data <- demo_data|> 
  mutate(ImporvementMethodology = if_else(IMP_FOCUS == 4, Other_ManFocus, as.character(IMP_FOCUS))
)


# Suppose the first two columns are A and B, and you want X to be the third column
demo_data <- demo_data %>% select(ORG_CTRY, GOV_LVL, FuncationalArea, POP_SIZE, InstitutionalizationLevel, ORG_SIZE, IMP_FOCUS, Other_ManFocus, ImporvementMethodology, everything())

# Remove extra column 
demo_data <- demo_data |> 
  select(-7:-8)
```

```{r}
#classify your experience 
#1- Academic/Researcher, 2- Practitioner in an LGO, 3- Consultant, 4- Other (Please specify)
names(demo_data)
table(demo_data$EXP)
table(demo_data$Other_EXP)


# to comibed other with the original column Functionla area:
demo_data <- demo_data|> 
  mutate(ExperienceRole = if_else(EXP == 4, Other_EXP, as.character(EXP))
)


# Remove extra column 
demo_data <- demo_data |> 
  select(-35:-36)
```

```{r}
#Role type  
#1- Top-level leader/ manager in the organization, 2- Member of the team that guided LP/LSS development and/or implementation, etc
names(demo_data)
table(demo_data$ROLE_TYP)
table(demo_data$Other_Role)


# to comibed other with the original column Functionla area:
demo_data <- demo_data|> 
  mutate(RoleType = if_else(ROLE_TYP == 6, Other_Role, as.character(ROLE_TYP))
)


# Suppose the first two columns are A and B, and you want X to be the third column
demo_data <- demo_data %>% select(ORG_CTRY, GOV_LVL, FuncationalArea, POP_SIZE, InstitutionalizationLevel, ORG_SIZE, ImporvementMethodology, everything())

# Remove extra column 
demo_data <- demo_data |> 
  select(-36:-37)
```


```{r}
#I still need to clean the tools and align the variables info inside 
#write.csv(demo_data, "demo_data_afterCleaning.csv")
```


# Taking all data to align with data for statistical model 
```{r}
#Total NAs per participant for dependent and independent variables/items   
names(demo_data)
# select aqll except tools 
demo_data_Non_tools <- demo_data |> 
  select(c(1:7, 34:39))

demo_data_Non_tools <- demo_data_Non_tools %>% 
  mutate(total_na1 = rowSums(is.na(demo_data_Non_tools)))

NAs_per_part1 <- ggplot(demo_data_Non_tools, aes(total_na1)) +
  geom_bar(fill = "maroon") + # to fill bar color 
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -0.8) +
  scale_x_continuous(breaks = seq(min(demo_data_Non_tools$total_na1, na.rm = TRUE), 
                                  max(demo_data_Non_tools$total_na1, na.rm = TRUE), 
                                  by = 2)) +
  labs(x = "Total Missing Responses", 
       y = "Participant", 
       title = "Missing Responses per Participant") +
  ylim(0, 100) 
NAs_per_part1

# filter data by selecting those with less than 11 missing 
demo_data_Non_tools <- demo_data_Non_tools %>% 
  filter(total_na1 <= 5)

# remove defense: one of the funcationla area 
demo_data_Non_tools <- demo_data_Non_tools |> 
  filter(!FuncationalArea %in%
                            "Defense")
```

# Align names in each variabel 
```{r}
#names(demo_data_Non_tools)
table(demo_data_Non_tools$FuncationalArea)

# Reassign numeric to text 
demo_data_Non_tools<- demo_data_Non_tools |> 
  mutate(FuncationalArea = case_when(
    FuncationalArea == "1" ~ "Social Services",
    FuncationalArea == "10" ~ "Healthcare",
    FuncationalArea == "3"~ "Transportation",
    FuncationalArea == "4" ~ "Finance and taxation",
    FuncationalArea == "5" ~ "School and education",
    FuncationalArea == "7" ~ "Court", 
    FuncationalArea == "9" ~ "Administration services", 
    TRUE ~ FuncationalArea  # Keeps the original FuncationalArea if none of the above conditions are met
  ))

# Clean and organizations Funcationla area
demo_data_Non_tools <- demo_data_Non_tools %>%
  mutate(FuncationalArea = case_when(
    FuncationalArea %in% c("Industrial Insurance / Labor Standards / Safety & Health", "All agencies in the organization", "All State Services", "Water/Sewer, Legal, Building &Housing, Finance") ~ "All/Multiple Funcational Area",
    FuncationalArea %in% c("Environmental Management", "Environmental Protection", "Environmental") ~ "Environmental Protection",
    FuncationalArea %in% c("Public Health") ~ "Healthcare",
    TRUE ~ FuncationalArea  # Keeps the original FuncationalArea if none of the above conditions are met
  ))
```


# Institutionalization Level
```{r}
table(demo_data_Non_tools$InstitutionalizationLevel)

demo_data_Non_tools <- demo_data_Non_tools |> 
  mutate(InstitutionalizationLevel = case_when(
    InstitutionalizationLevel == "1" ~ "Within a single sub-unit of the organization",
    InstitutionalizationLevel == "2" ~ "Within multiple sub-units of the organization",
    InstitutionalizationLevel == "3" ~ "Across all sub-units of the organization",
    TRUE ~ InstitutionalizationLevel
  ))

demo_data_Non_tools <- demo_data_Non_tools |> 
  mutate(InstitutionalizationLevel = case_when(
    InstitutionalizationLevel %in% c("Multiple departments within and Agency of the County") ~ "Within multiple sub-units of the organization",
    TRUE ~ InstitutionalizationLevel  # Keeps the original InstitutionalizationLevel if none of the above conditions are met
  ))
```

# ORG_SIZE
```{r}
table(demo_data_Non_tools$ORG_SIZE)

demo_data_Non_tools <- demo_data_Non_tools |> 
  mutate(ORG_SIZE = case_when(
    ORG_SIZE == "1" ~ "Less than 100 employees",
    ORG_SIZE == "2" ~ "Between 101 to 250 employees",
    ORG_SIZE == "3" ~ "Between 251 to 500 employees",
    ORG_SIZE == "4" ~ "Between 501 to 750 employees",
    ORG_SIZE == "5" ~ "Between 751 to 1,000 employees",
    ORG_SIZE == "6" ~ "More than 1,000 employees",
    TRUE ~ InstitutionalizationLevel
  ))
```

# Improvement Methodology 
```{r}
table(demo_data_Non_tools$ImporvementMethodology)

demo_data_Non_tools <- demo_data_Non_tools |> 
  mutate(ImporvementMethodology = case_when(
    ImporvementMethodology == "1" ~ "Lean Production", 
    ImporvementMethodology == "2" ~"Lean Six Sigma",
    ImporvementMethodology == "3" ~ "Lean Production with other general improvement
methodology",

TRUE ~ImporvementMethodology
  ))

demo_data_Non_tools <- demo_data_Non_tools |> 
  mutate(ImporvementMethodology = case_when(
    ImporvementMethodology %in% c("DMAIC") ~ "Six Sigma",
    ImporvementMethodology %in% c("Lean Services")~ "Lean Production", 
    
    TRUE ~ ImporvementMethodology
  ))

```

# Percentaqe Implemented 
```{r}
table(demo_data_Non_tools$IMPL_PCT)

demo_data_Non_tools <- demo_data_Non_tools |> 
  mutate(
    IMPL_PCT = as.character(IMPL_PCT),# Convert IMPL_PCT to character as it did not run without it 
    IMPL_PCT = case_when(
    IMPL_PCT == "1" ~ "0-20%", 
    IMPL_PCT == "2" ~"21%-40%",
    IMPL_PCT == "3" ~"41%-60%",
    IMPL_PCT == "4" ~"61%-80%",
    IMPL_PCT == "5" ~ "81%-100%",

TRUE ~IMPL_PCT
  ))
```

#years of experience 
```{r}
table(demo_data_Non_tools$EXP_YRS)
demo_data_Non_tools <- demo_data_Non_tools |> 
  mutate(
    EXP_YRS = as.character(EXP_YRS),# Convert IMPL_PCT to character as it did not run without it 
    EXP_YRS = case_when(
    EXP_YRS == "1" ~ "Less than a year", 
    EXP_YRS == "2" ~"1-5 years",
    EXP_YRS == "3" ~"6-10 years",
    EXP_YRS == "4" ~"11-15 years",
    EXP_YRS == "5" ~ "16-20 years",
    EXP_YRS == "6" ~ "More than 20 years",

TRUE ~EXP_YRS
  ))
```

# How long ago was this LP/LSS completed
```{r}
table(demo_data_Non_tools$COMP_AGO)
demo_data_Non_tools <- demo_data_Non_tools |> 
  mutate(
    COMP_AGO = as.character(COMP_AGO),# Convert IMPL_PCT to character as it did not run without it 
    COMP_AGO = case_when(
    COMP_AGO == "1" ~ "Less than 6 months ago", 
    COMP_AGO == "2" ~"Between 6 and 12 months ago",
    COMP_AGO == "3" ~"Between 12 and 18 months ago",
    COMP_AGO == "4" ~"Between 12 and 18 months ago",
    COMP_AGO == "5" ~ "More than 24 months ago",

TRUE ~COMP_AGO
  ))
```

#How long has the organization been working on LP/LSS
```{r}
table(demo_data_Non_tools$ORG_EXP)
demo_data_Non_tools <- demo_data_Non_tools |> 
  mutate(
    ORG_EXP = as.character(ORG_EXP),# Convert IMPL_PCT to character as it did not run without it 
    ORG_EXP = case_when(
    ORG_EXP == "1" ~ "Less than 1 year", 
    ORG_EXP == "2" ~"Between 1 and 2 years",
    ORG_EXP == "3" ~"Between 3 and 4 years",
    ORG_EXP == "4" ~"Between 5 and 6 years",
    ORG_EXP == "5" ~ "More than 6 years",

TRUE ~ORG_EXP
  ))
```

# How would you classify your experience with LP/LSS?
```{r}
table(demo_data_Non_tools$ExperienceRole)
demo_data_Non_tools <- demo_data_Non_tools |> 
  mutate(
    ExperienceRole = as.character(ExperienceRole),# Convert IMPL_PCT to character as it did not run without it 
    ExperienceRole = case_when(
    ExperienceRole == "1" ~ "Academic/Researcher", 
    ExperienceRole == "2" ~"Practitioner in an LGO",
    ExperienceRole == "3" ~"Consultant",

TRUE ~ExperienceRole
  ))

demo_data_Non_tools <- demo_data_Non_tools |> 
  mutate(ExperienceRole = case_when(
    ExperienceRole %in% c("Internal Consultant", "Manager of a Lean Office") ~ "Practitioner in an LGO",
    
    TRUE ~ExperienceRole
  ))



```


# best describes your role (or roles) in LP/LSS development
```{r}
table(demo_data_Non_tools$RoleType)
demo_data_Non_tools <- demo_data_Non_tools |> 
  mutate(
    RoleType = as.character(RoleType),# Convert IMPL_PCT to character as it did not run without it 
    RoleType = case_when(
    RoleType == "1" ~ "Top-level leader/ manager in the organization", 
    RoleType == "2" ~"Member of the team that guided LP/LSS development and/or implementation",
    RoleType == "3" ~"Consultant within the department or from other departments/units within the organization",
    RoleType == "4" ~"External consultant",
    RoleType == "5" ~"Process owner",
    
TRUE ~RoleType
  ))

demo_data_Non_tools <- demo_data_Non_tools |> 
  mutate(RoleType = case_when(
    RoleType %in% c("Lean Facilitator within the unit", "Implement LP/LSS in directly assigned work task and life") ~ "Consultant within the department or from other departments/units within the organization",
    RoleType %in% c("Organizational Learning & Development") ~ "Top-level leader/ manager in the organization",
    
    TRUE ~RoleType
  ))

```

# To Print data: descriptive summary fo all demographic varibales except POP-SIZE 
```{r}
demo_data_Non_tools <- demo_data_Non_tools |> 
  select(c(-14:-22))

names(demo_data_Non_tools)
library(crosstable)
# All observation 
demo_data_Non_tools_summary <- crosstable(demo_data_Non_tools)
demo_data_Non_tools_summary
write.csv(demo_data_Non_tools_summary, "descriptive_demographic_WO_POP_SIZE.csv")
```


# Population size 
```{r}
#write.csv(demo_data_Non_tools, "demographic_Data.csv")
# Note: here i donwload the data after clean all variables and ajust Pop-SIZE mannually and then reupload the data 

New_demo_data_Non_tools <- read.csv("demographic_Data.csv")
New_demo_data_Non_tools <- New_demo_data_Non_tools |> 
  select(-1)
names(New_demo_data_Non_tools)

# Convert cells with not defined NAs to NAs 
New_demo_data_Non_tools[New_demo_data_Non_tools == 'NA'] <- NA
# convet empty cells to NAs
New_demo_data_Non_tools[New_demo_data_Non_tools == ''] <- NA
# To remove comma from the observations 
New_demo_data_Non_tools$POP_SIZE <- gsub(",", "", New_demo_data_Non_tools$POP_SIZE)

New_demo_data_Non_tools$POP_SIZE <- as.numeric(New_demo_data_Non_tools$POP_SIZE)


New_demo_data_Non_tools <- New_demo_data_Non_tools |> 
  mutate(POP_SIZE_New = case_when(
    POP_SIZE <= 100000 ~ "<=100K",
    POP_SIZE >100000 & POP_SIZE <= 500000 ~ "100K-500K",
    POP_SIZE > 500000 & POP_SIZE <= 1000000 ~ "500K-1M",
    POP_SIZE>1000000 ~ ">1M"
  ))























```



# Tools
```{r}
names(demo_data)

Tools_LPLss <- demo_data |> 
  select(8:33)

Tools_LPLss <- Tools_LPLss %>% 
  mutate(total_na1 = rowSums(is.na(Tools_LPLss)))

NAs_per_part1 <- ggplot(Tools_LPLss, aes(total_na1)) +
  geom_bar(fill = "maroon") + # to fill bar color 
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -0.8) +
  scale_x_continuous(breaks = seq(min(Tools_LPLss$total_na1, na.rm = TRUE), 
                                  max(Tools_LPLss$total_na1, na.rm = TRUE), 
                                  by = 2)) +
  labs(x = "Total Missing Responses", 
       y = "Participant", 
       title = "Missing Responses per Participant") +
  ylim(0, 100) 
NAs_per_part1

Tools_LPLss <- Tools_LPLss %>% 
  filter(total_na1 <= 24)

sum(table(Tools_LPLss$VSM))

table(Tools_LPLss$Other_Tools)

# Change texts in column other_tools to other

Tools_LPLss <- Tools_LPLss %>%
  mutate(Other_Tools = case_when(
    Other_Tools %in% c("Kaizen every day, stand-up meetings") ~ "1",
    Other_Tools %in% c("Customer Journey Mapping; Usability Studies") ~ "1",
    Other_Tools %in% c("PICK Chart") ~ "1",
    Other_Tools %in% c("Process Instructions Development") ~ "1",
    Other_Tools %in% c("process walks, data collection plan, implementation plan, etc.") ~ "1",
    Other_Tools %in% c("SIPOC") ~ "1",
    Other_Tools %in% c("statistical analysis") ~ "1",
    Other_Tools %in% c("Obeya") ~ "1",
    
    TRUE ~ Other_Tools  # Keeps the original FuncationalArea if none of the above conditions are met
  ))

Tools_LPLss <- Tools_LPLss |> 
  select(-27)

```

# Remove duplicte columns 
```{r}
# remove duplicate column: Project_Charter_2
sum(!is.na(Tools_LPLss$Project_Charter))
sum(!is.na(Tools_LPLss$Project_Charter_2))
names(Tools_LPLss)

# Assuming you have a dataframe named df with columns Project_Charter and Project_Charter_2

# Combine the two columns as described
Tools_LPLss$Project_Charter <- ifelse(is.na(Tools_LPLss$Project_Charter) & Tools_LPLss$Project_Charter_2 == 1, 
                             1, 
                             Tools_LPLss$Project_Charter)

# This will leave Project_Charter as is if it's 1, 
# but replace NA with 1 if Project_Charter_2 is 1.


# # remove duplicate column: Benchmarking_2
sum(!is.na(Tools_LPLss$Benchmarking_2))
sum(!is.na(Tools_LPLss$Benchmarking))

#bench <- Tools_LPLss |> 
  select(11, 19)

Tools_LPLss$Benchmarking <- ifelse(is.na(Tools_LPLss$Benchmarking) & Tools_LPLss$Benchmarking_2 == 1, 
                             1, 
                             Tools_LPLss$Benchmarking)
Tools_LPLss <- Tools_LPLss |> 
  select(c(-18:-19))
```

# Plot tools 
```{r}
names(Tools_LPLss)
Tools_LPLss<- Tools_LPLss |> 
  rename(
    VSM = VSM,
    "5S" = FIVE_S,
    "Visual Management" = Visual_Mgmt,
    "5Whys"= Five_Whys,
    "Process Flow Chart" = PFC,
    "Kaizen Events" = KEs,
    "Cause and Effect Diagram/Analysis" = CEDA,
    "Mistake Proofing" = Mistake_Proofing, 
    "Project Charter" = Project_Charter,
    "Gemba" = Gemba, 
    Benchmarking = Benchmarking,
    "Spaghetti Diagram" = Spaghetti_Diagram,
    "Statistical Process Control" = SPC, 
    "Check Sheet Kanban"= CS_Kanban, 
    "Critical to Quality" = CTQ,
    "One-Piece Flow" = OPF,
    "Quality Function Deployment"= QFD,
    "Cycle Time Analysis" = CTA,
    "Policy Deployment"= PD,
    A3 = A3,
    Brainstorming = Brainstorming.,
    "Lead Time Analysis"= LTA,
    "Single-Minute Exchange of Dies" =SMED, 
    Other_Tools = Other_Tools
    
    
  )

library(tidyr)
library(ggplot2)
library(dplyr)

# Assume Tools_LPLss is your dataframe
# Replace empty strings with NA in the Other_Tools column
Tools_LPLss$Other_Tools[Tools_LPLss$Other_Tools == ""] <- NA

# Convert all 'NA' strings to actual NA values
Tools_LPLss[Tools_LPLss == 'NA'] <- NA

# Reshape the data from wide to long format
long_data <- pivot_longer(Tools_LPLss, cols = everything(), names_to = "Variable", values_to = "Value")

# Count the number of non-NA values for each variable and sort it
long_data_sorted <- long_data %>%
  filter(!is.na(Value)) %>% # Filter out NA values
  group_by(Variable) %>%
  summarise(Count = n()) %>%
  ungroup() %>%
  arrange(desc(Count)) %>%
  mutate(Variable = factor(Variable, levels = Variable))




# Now use ggplot2 to create the plot with sorted bars
ggplot(long_data_sorted, aes(x = Variable, y = Count, fill = Variable)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "LP/LSS Tools", x = "Tools", y = "Frequency") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

