---
title: "EvaluteMissingPercentage.Branch"
author: "Mohammed Alrezq"
date: "2023-07-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The following is to check which sample size tp pick and has 10% missing per variable 
```{r}
#with missing 4 or less 
data.CompletedCases4 <- factor_data |> 
  filter(total_na <= 4) |> 
  select(-total_na)
# Plot variables for missing% from the data.CompletedCases4
CompletedCases.plot4 <- gg_miss_var(data.CompletedCases4, show_pct = TRUE) +
  theme(plot.background = element_rect(fill = "white")) +
  plotTheme() + 
  labs(title = "Participants with 4 missing questions")
# with missing 6 or less 
data.CompletedCases6 <- factor_data |> 
  filter(total_na <= 6) |> 
  select(-total_na)
# Plot variables for missing% from the data.CompletedCases6
CompletedCases.plot6 <- gg_miss_var(data.CompletedCases6, show_pct = TRUE) +
  theme(plot.background = element_rect(fill = "white")) +
  plotTheme() + 
  labs(title = "Participants with 6 missing questions")
# with missing 9 or less 
data.CompletedCases9 <- factor_data |> 
  filter(total_na <= 9) |> 
  select(-total_na)
# Plot variables for missing% from the data.CompletedCases9
CompletedCases.plot9 <- gg_miss_var(data.CompletedCases9, show_pct = TRUE) +
  theme(plot.background = element_rect(fill = "white")) +
  plotTheme() + 
  labs(title = "PParticipants with 9 missing questions")
# with missing 10 or less 
data.CompletedCases10 <- factor_data |> 
  filter(total_na <= 10) |> 
  select(-total_na)
# Plot variables for missing% from the data.CompletedCases10
CompletedCases.plot10 <- gg_miss_var(data.CompletedCases10, show_pct = TRUE) +
  theme(plot.background = element_rect(fill = "white")) +
  plotTheme()+
  labs(title = "Participants with 10 missing questions")
# with missing 19 or less 
data.CompletedCases19 <- factor_data |> 
  filter(total_na <= 19) |> 
  select(-total_na)
# Plot variables for missing% from the data.CompletedCases19
CompletedCases.plot19 <- gg_miss_var(data.CompletedCases19, show_pct = TRUE) +
  theme(plot.background = element_rect(fill = "white")) +
  plotTheme() +
  labs(title = "Participants with 19 missing questions") 
```

# Imputation using MICE Package 

## Plot using md.pattern to check the pattern of missingness  
```{r}
install.packages("VIM")# this package is to plot the pattern in missing dataset 
library(VIM)
install.packages("MICE")
library(mice)

# delete response ID and total_na columns 
factor_data_MICE <- factor_data |> 
  select(c(-Response.ID, -total_na))

# missing per variabels 
md.pattern(factor_data_MICE, rotate.names = TRUE)



```

## The following package can also be used to visulize missing data 
- Package name: DataExplorer 
```{r}
#install.packages("DataExplorer")
library(DataExplorer)
plot_missing(factor_data_MICE, 
             missing_only = TRUE) +
  plotTheme()

```

## Imputation using single methdo: mean
```{r}
# The good things about MICE imputtion using mean is that it imputes same value from exisitng data, meaning that plusable vlaues are ordinal not decimal 
imputation.1 <- mice(factor_data_MICE,
                     method = "mean", # Replace by mean of the other values 
                     m = 1, # No. of multiple imputation 
                     maxit = 1) # max. iteration. it is good for convergence 


# to check/see imputed values for each missing cell in each variable
imputation.1$imp

# To see the completed dataset after imputation 
complete(imputation.1, action = 1) # action is the first dataset, and it is only 1 dataset as we have one # iteration



density(imputation.1)

str(imputation.1)


library(VIM)
aggr_plot <- aggr(factor_data_MICE, 
                  col=c('lightblue','lightgreen'), 
                  numbers=TRUE, 
                  sortVars=TRUE, 
                  labels=names(factor_data_MICE), 
                  cex.axis=.3, 
                  gap=3, 
                  ylab=c("Histogram of missing data","Pattern"))+
  plotTheme()

A <- aggr(factor_data_MICE, plot = FALSE)
plot(A, numbers = TRUE, prop = FALSE)
```


# check if the data fopr missing is missing at MCAR
- [This video explain well about MCAR](https://www.youtube.com/watch?v=h9CzBtLpt_8&ab_channel=StatswithMia)

```{r}
# install the package if you haven't already
install.packages("naniar")

# load the package
library(naniar)

# use the test_mcar function on your data
result <- mcar_test(factor_data_MICE)

# print the result
print(result)
# A tibble: 1 × 4
#   statistic    df  p.value missing.patterns
#       <dbl> <dbl>    <dbl>            <int>
# 1      548.   349 4.51e-11               12
# # the above results from mcar_test suggested that the missininges is not related to MCAR



```

## Another package to imputation using regression 
```{r}
install.packages("Hmisc")
library(Hmisc)


```


to use kinti package for table 
```{r}
install.packages("knitr")
library(knitr)

summ <- summary(data)

knitr::kable(summ)

```

```{r}
plotTheme <- function() {
  theme(
    plot.title = element_text(size = 14, family = "sans", face = "plain", hjust = 0),
    plot.subtitle=element_text(size = 11, family = "sans", hjust = 0),
    plot.caption=element_text(size = 10, family = "sans", face = "italic", hjust = 0), 
    axis.title.x = element_text(size = 10, family = "sans", face = "plain", hjust = 1, vjust = -0.5),
    axis.title.y = element_text(size = 10, family = "sans", face = "plain", hjust = 1, vjust = 1),
    axis.text = element_text(size = 9, family = "sans", face = "plain"),
    panel.background = element_blank(),
    panel.grid.minor = element_line(colour = "gray"),
    panel.grid.major = element_line(colour = "gray"),
    axis.ticks = element_blank(),
    legend.title = element_text(size = 10, family = "sans"),
    legend.text = element_text(size = 9, family = "sans"),
    axis.line = element_blank()
  )
}

```

# Factor analysis Steps:
## 1.2 Load data 
```{r}
data <- read.csv("mean_data.csv")

# Remove Response.ID and col X from the data 
data <- data |> 
  select(c(-Response.ID, -X))

# Convert all integer columns to numeric
data <- data |> 
  mutate_if(is.integer, as.numeric)
str(data)
dim(data)
```

# 2. Preliminary Steps 
## 2.1 Descriptive statistics 
```{r}
describe(data)
```

## 2.2 Check normality plot 
```{r}
# Function to plot a histogram for a variable
plot_histogram <- function(var) {
  ggplot(data, aes_string(var)) +
    geom_histogram(binwidth = 0.5, color = "black", fill = "lightblue") +
    labs(x = var, y = "Frequency") +
    theme_minimal() +
    plotTheme()
}

# Get the names of the variables
vars <- names(data)

# Plot a histogram for each variable
plots <- lapply(vars, plot_histogram)

# Print the plots
grid.arrange(grobs = plots, ncol = 4)


# another way to check normality 
library(lattice)
cat(names(data), sep = " + ")
histogram(~IMPL_SUCC + CULT_INTG + BUDG_COMP + ORG_PERF_IMP + EMPL_SKL_IMP + DEPLOY + ACT_INTG + BUSI_IMP + TASK_COMP + STAKEH_ACC + EMPL_SAT + EMPL_BEL_IMP + DEC_IMP + LD_SUP + ORG_CUL + EMP_ATT + RES_AVL + LD_INV + PRJ_SEL + DEF_MEA + RES_CHG + LD_STL + ORG_CHR + DAT_QLT + EMP_ENG + TOOL_AP + CONTX_CUS + EFT_REQ + SCOPE + INSTIT + TRA_EDUC + WKF_SKL + EXP_CON + EXT_CON + UND_CNG + COM_KNW + PAT_RES + GOAL_CL + CUS_IDEF + GOAL_AL + SYS_COM + TEAM_ST, data = data)# this plot might not be good as imputed value shown on their own bar
```

# Multivariate normality 
- Run mardia's multivariate normality test 
```{r}
mardia(data)
# check what test used for normality check 
# In our case, kurtosis = -3.2 (P < 0.05). At the start and the end, the dots are away from the line. Thus, the data are not normally distributed at multivariate level. Our extraction method PAF can deal with this non-normality.
```


# data factorability 
## 1. Kaiser-Meyer-Olkin (KMO)
 
```{r}
KMO(data)
#Overall MSA =  0.86 whic h is good 
```

## 2 Bartlet’s test of sphericity
```{r}
cortest.bartlett(data)
# $p.value = 5.660815e-161, which means items are correlated 
```

# Determine No. of factor 
# 1 scree plot and eginvalue 
```{r}
scree(data)

# based on the plot
#  FA = 5 factor 
#  PC  = 8 
```

# 2 parallel analysis 
```{r}
par <- fa.parallel(data, fm = "pa", fa = "fa")
vss(data)
```


# Run factor analysis with 5 factors
```{r}
Prn_ax_fac <- fa(data, nfactors = 5, fm = "pa", rotate = "oblimin")
print(Prn_ax_fac , cut = .3, digits = 3)

fa.diagram(Prn_ax_fac)

```

# Run factor analysis with 10 factors
```{r}
Prn_ax_fac_10 <- fa(data, nfactors = 10, fm = "pa", rotate = "oblimin")
print(Prn_ax_fac_10 , cut = .3, digits = 3)

fa.diagram(Prn_ax_fac_10)

loading_10 <- Prn_ax_fac_10$loadings
#write.csv(loading_10, "fact_loading_10fac.csv")

# make no sense 

```

# Run factor analysis with 6 factors
```{r}
Prn_ax_fac_6 <- fa(data, nfactors = 6, fm = "pa", rotate = "oblimin")
print(Prn_ax_fac_10 , cut = .3, digits = 3)

fa.diagram(Prn_ax_fac_6)

loading_6 <- Prn_ax_fac_6$loadings
#write.csv(loading_6, "fact_loading_6fac.csv")


```

# another way: Run FA fro DVs and IVs seperatly 

```{r}
names(data)
# dependent variables 
DVs <- data |> 
  select(c(1:13))
# independent variables 
IVs <- data |> 
  select(c(14:42))
```

# DVs check normality 
```{r}
# Multivariate normality 
mardia(DVs) # not normal 

## Kaiser-Meyer-Olkin (KMO)
KMO(DVs) # overall MSA =0.9, threshold = 0.7

#Bartlet’s test of sphericity

cortest.bartlett(DVs)
#p.value = 7.086821e-94, which means items are correlated 
```

# No. of factor for DVs 
```{r}
scree(DVs)
```

# Run factor analysis with 6 factors
```{r}
Prn_ax_fac_DV_2 <- fa(DVs, nfactors = 4, fm = "pa", rotate = "oblimin")
print(Prn_ax_fac_DV_2 , cut = .3, digits = 2)

fa.diagram(Prn_ax_fac_DV_2)
```

# Independent variables 
```{r}
# Multivariate normality 
mardia(IVs) # not normal 

## Kaiser-Meyer-Olkin (KMO)
KMO(IVs) # overall MSA =0.9, threshold = 0.7

#Bartlet’s test of sphericity

cortest.bartlett(IVs)
#p.value = 7.086821e-94, which means items are correlated 
```

# No. of factor for IVs 
```{r}
scree(IVs)
```

# Run factor analysis with 4 factors
```{r}
Prn_ax_fac_IV <- fa(IVs, nfactors = 4, fm = "pa", rotate = "oblimin")
print(Prn_ax_fac_IV , cut = .3, digits = 2)

fa.diagram(Prn_ax_fac_IV)
```

# Run factor analysis with 5 factors
```{r}
Prn_ax_fac_IV_5 <- fa(IVs, nfactors = 5, fm = "pa", rotate = "oblimin")
print(Prn_ax_fac_IV_5 , cut = .3, digits = 2)

# eginvalues: these are the eginvalue from scree. often above or equal 1 
Prn_ax_fac_IV_5$e.values
```

# Principle component analysis 
```{r}
pc.IVs <- principal(IVs, nfactors = 7, rotate = "oblimin")# n.factor based on scree
print(pc.IVs, cut = 0.3, digits = 2)
```

