---
title: "EvaluteMissingPercentage.Branch"
author: "Mohammed Alrezq"
date: "2023-07-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The following is to check which sample size tp pick and has 10% missing per variable 
```{r}
#with missing 4 or less 
data.CompletedCases4 <- factor_data |> 
  filter(total_na <= 4) |> 
  select(-total_na)
# Plot variables for missing% from the data.CompletedCases4
CompletedCases.plot4 <- gg_miss_var(data.CompletedCases4, show_pct = TRUE) +
  theme(plot.background = element_rect(fill = "white")) +
  plotTheme() + 
  labs(title = "Participants with 4 missing questions")
# with missing 6 or less 
data.CompletedCases6 <- factor_data |> 
  filter(total_na <= 6) |> 
  select(-total_na)
# Plot variables for missing% from the data.CompletedCases6
CompletedCases.plot6 <- gg_miss_var(data.CompletedCases6, show_pct = TRUE) +
  theme(plot.background = element_rect(fill = "white")) +
  plotTheme() + 
  labs(title = "Participants with 6 missing questions")
# with missing 9 or less 
data.CompletedCases9 <- factor_data |> 
  filter(total_na <= 9) |> 
  select(-total_na)
# Plot variables for missing% from the data.CompletedCases9
CompletedCases.plot9 <- gg_miss_var(data.CompletedCases9, show_pct = TRUE) +
  theme(plot.background = element_rect(fill = "white")) +
  plotTheme() + 
  labs(title = "PParticipants with 9 missing questions")
# with missing 10 or less 
data.CompletedCases10 <- factor_data |> 
  filter(total_na <= 10) |> 
  select(-total_na)
# Plot variables for missing% from the data.CompletedCases10
CompletedCases.plot10 <- gg_miss_var(data.CompletedCases10, show_pct = TRUE) +
  theme(plot.background = element_rect(fill = "white")) +
  plotTheme()+
  labs(title = "Participants with 10 missing questions")
# with missing 19 or less 
data.CompletedCases19 <- factor_data |> 
  filter(total_na <= 19) |> 
  select(-total_na)
# Plot variables for missing% from the data.CompletedCases19
CompletedCases.plot19 <- gg_miss_var(data.CompletedCases19, show_pct = TRUE) +
  theme(plot.background = element_rect(fill = "white")) +
  plotTheme() +
  labs(title = "Participants with 19 missing questions") 
```

# Imputation using MICE Package 

## Plot using md.pattern to check the pattern of missingness  
```{r}
install.packages("VIM")# this package is to plot the pattern in missing dataset 
library(VIM)
install.packages("MICE")
library(mice)

# delete response ID and total_na columns 
factor_data_MICE <- factor_data |> 
  select(c(-Response.ID, -total_na))

# missing per variabels 
md.pattern(factor_data_MICE, rotate.names = TRUE)



```

## The following package can also be used to visulize missing data 
- Package name: DataExplorer 
```{r}
#install.packages("DataExplorer")
library(DataExplorer)
plot_missing(factor_data_MICE, 
             missing_only = TRUE) +
  plotTheme()

```

## Imputation using single methdo: mean
```{r}
# The good things about MICE imputtion using mean is that it imputes same value from exisitng data, meaning that plusable vlaues are ordinal not decimal 
imputation.1 <- mice(factor_data_MICE,
                     method = "mean", # Replace by mean of the other values 
                     m = 1, # No. of multiple imputation 
                     maxit = 1) # max. iteration. it is good for convergence 


# to check/see imputed values for each missing cell in each variable
imputation.1$imp

# To see the completed dataset after imputation 
complete(imputation.1, action = 1) # action is the first dataset, and it is only 1 dataset as we have one # iteration



density(imputation.1)

str(imputation.1)


library(VIM)
aggr_plot <- aggr(factor_data_MICE, 
                  col=c('lightblue','lightgreen'), 
                  numbers=TRUE, 
                  sortVars=TRUE, 
                  labels=names(factor_data_MICE), 
                  cex.axis=.3, 
                  gap=3, 
                  ylab=c("Histogram of missing data","Pattern"))+
  plotTheme()

A <- aggr(factor_data_MICE, plot = FALSE)
plot(A, numbers = TRUE, prop = FALSE)
```


# check if the data fopr missing is missing at MCAR
- [This video explain well about MCAR](https://www.youtube.com/watch?v=h9CzBtLpt_8&ab_channel=StatswithMia)

```{r}
# install the package if you haven't already
install.packages("naniar")

# load the package
library(naniar)

# use the test_mcar function on your data
result <- mcar_test(factor_data_MICE)

# print the result
print(result)
# A tibble: 1 × 4
#   statistic    df  p.value missing.patterns
#       <dbl> <dbl>    <dbl>            <int>
# 1      548.   349 4.51e-11               12
# # the above results from mcar_test suggested that the missininges is not related to MCAR



```

## Another package to imputation using regression 
```{r}
install.packages("Hmisc")
library(Hmisc)


```

